Hardware-Accelerated Rendering of Intersecting Volumes using Boolean Operations

### 2013-06-24 paper.tex
Put caption of listing below it
Diagram for volume representations
Listing for Boolean AND implementation

For many years the computing community has created software to visualize data
from medical and scientific equipment.  Normally the data is returned as a
three-dimensional grid of values.  This grid is called a volume, and the
individual values are named voxels, similar to how the elements of a picture are
called pixels.

@@

\lstset{language=XML}
\lstset{markfirstintag=true}
\lstset{showstringspaces=false}
\lstset{caption=XML of second pass for first approach.}
\lstset{label=listing:xml-of-second-pass}
\begin{lstlisting}[frame=single,float]
<cull faces="front" />
<framebuffer>
  <attachment usage="color" link="coords-texture-1" />
  <use program="primary-coords-program">
    <group id="intersection-group">
      <booleanAnd of="cube-1 cube-2" />
    </group>
  </use>
</framebuffer>
<framebuffer>
  <attachment usage="color" link="coords-texture-2" />
  <use program="secondary-coords-program">
    <instance of="intersection-group" />
  </use>
</framebuffer>
<cull faces="back" />
<framebuffer>
  <attachment usage="color" link="results-texture" />
  <use program="second-pass-program">
    <uniform name="CoordsTexture1" link="coords-texture-1" />
    <uniform name="CoordsTexture2" link="coords-texture-2" />
    <uniform name="VolumeTexture1" link="volume-texture-1" />
    <uniform name="VolumeTexture2" link="volume-texture-2" />
    <uniform name="MVPMatrix" usage="modelviewprojection" />
    <instance of="intersection-group" />
  </use>
</framebuffer>
\end{lstlisting}

The Boolean operations required by the two approaches were implemented as two
separate nodes.  Currently both nodes take two cubes, which are assumed to be
axis-aligned.  {\tt BooleanAndNode} finds the intersection of the two cubes by
finding the minimums and maximums of the cubes in world space.  {\tt
BooleanXorNode} does the same, but then uses a variation of the well-known
Sutherland-Hodgman polyline-clipping algorithm to break up the cubes into pieces
around the intersection, as shown in figure \ref{clipping}.  Both nodes use the
{\tt NodeListener} interface to only update their geometry when the cubes move.

>>>>

For many years the computing community has created software to visualize data
from medical and scientific equipment.  Normally the data is returned as a
three-dimensional grid of values.  This grid is called a volume, and the
individual values are named voxels, similar to how the elements of a picture are
called pixels.  Figure \ref{representations} shows two different representations
of a volume.

\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{representations.pdf}
\caption{Representations of a volume.  One can think of a volume as a 3D grid of
voxels or as a stack of 2D images.}
\label{representations}
\end{figure}

@@

\lstset{language=XML}
\lstset{markfirstintag=true}
\lstset{showstringspaces=false}
\lstset{caption=XML of second pass for first approach.}
\lstset{captionpos=b}
\lstset{label=listing:xml-of-second-pass}
\begin{lstlisting}[frame=single,float]
<cull faces="front" />
<framebuffer>
  <attachment usage="color" link="coords-texture-1" />
  <use program="primary-coords-program">
    <group id="intersection-group">
      <booleanAnd of="cube-1 cube-2" />
    </group>
  </use>
</framebuffer>
<framebuffer>
  <attachment usage="color" link="coords-texture-2" />
  <use program="secondary-coords-program">
    <instance of="intersection-group" />
  </use>
</framebuffer>
<cull faces="back" />
<framebuffer>
  <attachment usage="color" link="results-texture" />
  <use program="second-pass-program">
    <uniform name="CoordsTexture1" link="coords-texture-1" />
    <uniform name="CoordsTexture2" link="coords-texture-2" />
    <uniform name="VolumeTexture1" link="volume-texture-1" />
    <uniform name="VolumeTexture2" link="volume-texture-2" />
    <uniform name="MVPMatrix" usage="modelviewprojection" />
    <instance of="intersection-group" />
  </use>
</framebuffer>
\end{lstlisting}

The Boolean operations required by the two approaches were implemented as two
separate nodes.  Currently both nodes take two cubes, which are assumed to be
axis-aligned.  {\tt BooleanAndNode} finds the intersection of the two cubes by
finding the minimums and maximums of the cubes in world space, as shown in
listing \ref{listing:boolean-and}.  {\tt BooleanXorNode} does the same, but then
uses a variation of the well-known Sutherland-Hodgman polyline-clipping
algorithm to break up the cubes into pieces around the intersection, as shown in
figure \ref{clipping}.  Both nodes use the {\tt NodeListener} interface to only
update their geometry when the cubes move.

\lstset{language=C++}
\lstset{caption=Boolean AND implementation.}
\lstset{captionpos=b}
\lstset{label=listing:boolean-and}
\begin{lstlisting}[frame=single,float]
Vec4 min = max(c1.min, c2.min);
Vec4 max = min(c1.max, c2.max);
\end{lstlisting}
